---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data


BRFSS is an ongoing surveillance system designed to measure behavioural risk factors for the non-institutionalized adult population (18 years of age and older)
The Behavioral Risk Factor Surveillance System (BRFSS) is a collaborative project between all of the states in the United States (US) and participating US territories and the Centers for Disease Control and Prevention (CDC).
The BRFSS is administered and supported by CDC's Population Health Surveillance Branch, under the Division of Population Health at the National Center for Chronic Disease Prevention and Health Promotion.  

It's objective is to collect uniform, state-specific data on preventive health practices and risk behaviours that are linked to chronic diseases, injuries, and preventable infectious diseases that affect the adult population.

The data refers to adult citizens (18 years age and older) that gives insights about aspects of their health and all the factors that may involve healthy -and unhealthy practices.


Data collection:
the BRFSS is intented as an observational study with no form of interference, therefore all relationships between the different variables may indicate correlations.

The collected data for the survey stretches across all 50 states plus affiliated US territories including the District of Columbia, Guam, and the Commonwealth of Puerto Rico.
The survey is generalizble to the US population as the observations and variables are distributed across the individuals background like, ethicity, educational background, income level etc.

In conducting the BRFSS landline telephone survey, interviewers collect data from a randomly selected adult in a household. In conducting the cellular telephone version of the BRFSS questionnaire, interviewers collect data from an adult who participates by using a cellular telephone and resides in private residence or college housing.



* * *

## Part 2: Research questions

**Research quesion 1:**

**Research quesion 2:**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

